Taking Dyads Seriously

### 2018-10-02 paper/4_app.tex
minor mods, need to fix gibler

We apply AME to three recent IR studies: \citet{reiter:stam:2003, weeks:2012, gibler:2017} published in the \textit{American Political Science Review}. Each of these studies use relational data of state interactions and propose both dyadic, monadic, and structural explanations for the behavior of actors in the system. We demonstrate the capabilities of AME with reference to existing studies in order to highlight several features of our network based approach. First, we show that simply by using the AME framework scholars can better model the data generating process behind events of interest. Second, the results of AME estimation are interpretable alongside results using standard approaches. Third, through using this approach we can also quantify the degree to which first-, second-, and third-order dependencies are present.

@@

The visualization of the sender random effects highlights that the behavior of countries does not fully accord with predictions from the covariates specified by Reiter \& Stam. Specifically, we can see that countries such as Iraq, Israel, Egypt, and Iran are more likely to be involved in initiating or continuing conflict with other countries than the model would predict. Further, other countries such as Bhutan, Finland, and Turkmenistan are less likely to engage in conflict than the fixed effects in the model would suggest. None of these findings change the key conclusion from Reiter \& Stam's work, but by using the AME framework we are able to better understand the limitations of their model.

@@

There is also a striking difference between Weeks' original model and our estimation using AME in terms of capturing the data generating process. Specifically, the area under the ROC curve for AME and GLM models are 0.97 and 0.64, respectively, which is a dramatic difference given that both models use the same fixed effects. As with Reiter \& Stam, the divergence is a result of the GLM framework's inability to account for underlying structure generating the event of interest. To uncover this structure, we illustrate another beneficial aspect of the AME framework, which is the multiplicative random effects estimated through the LFM portion of the model -- $\mathbf{u}_{i}^{\top} \mathbf{D} \mathbf{v}_{j}$. These random effects account for higher order dependencies that manifest as a result of homophily and stochastic equivalence in dyadic data.

>>>>

We apply AME to three recent IR studies: \citet{reiter:stam:2003, weeks:2012, gibler:2017}. Each of these studies use relational data of state interactions and propose both dyadic, monadic, and structural explanations for the behavior of actors in the system. We demonstrate the capabilities of AME with reference to existing studies in order to highlight several features of our network based approach. First, we show that simply by using the AME framework scholars can better model the data generating process behind events of interest. Second, the results of AME estimation are interpretable alongside results using standard approaches. Third, through using this approach we can also quantify the degree to which first-, second-, and third-order dependencies are present.

@@

The visualization of the sender random effects highlights that the behavior of countries does not fully accord with predictions from the covariates specified by Reiter \& Stam. Specifically, we can see that countries such as Iraq, Israel, Egypt, and Iran are more likely to be involved in initiating or continuing conflict with other countries than the model would predict. Further, other countries such as Bhutan, Finland, and Turkmenistan are less likely to engage in conflict than the exogenous covariates in the model would suggest. None of these findings change the key conclusion from Reiter \& Stam's work, but by using the AME framework we are able to better understand the limitations of their model.

@@

There is also a striking difference between Weeks' original model and our estimation using AME in terms of capturing the data generating process. Specifically, the area under the ROC curve for AME and GLM models are 0.97 and 0.64, respectively, which is a dramatic difference given that both models use the same exogenous covariates. As with Reiter \& Stam, the divergence is a result of the GLM framework's inability to account for underlying structure generating the event of interest. To uncover this structure, we illustrate another beneficial aspect of the AME framework, which is the multiplicative random effects estimated through the LFM portion of the model -- $\mathbf{u}_{i}^{\top} \mathbf{D} \mathbf{v}_{j}$. These random effects account for higher order dependencies that manifest as a result of homophily and stochastic equivalence in dyadic data.
